---
title: "Hot Ones Episodes"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

The reason we chose this specific dataset is because a couple of us in the group have seen the show and thought it would be fun. It also seems like it would be engaging to present for the rest of the class.

## Questions

Q1 ) Is there any correlation between the scoville score of the sauce and the number of guests who did not finish trying all the sauces?
1.1) Find the season which has the most unfinished contestants
1.2) Compare that with the sauce score.



## Analysis plan

- For the first question, we plan on creating the variable "unfinished_count" which represents the number of guests that failed to try all the sauces for each season. Then we create another variable "average_scoville" which is the average of the variable "scoville" for each season. Then we can make a scatter plot in the first layer with "average_scovile" on the x-axis and "unfished_count" on the y-axis. We will also add a geom_smooth in the second layer to analyse the trend.
Here we calculate the "unfinished_count" variable from "episodes.csv" and the "average_scovile" variable from "sauces.csv". 

