---
title: "Proposal title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

## Questions

The two questions you want to answer.

## Analysis plan

- For the first question, we plan on creating the variable "unfinished_count" which represents the number of celebrities that failed to try all the sauces for each season. Then we create another variable "average_scoville" which is the average of the variable "scoville" for each season. Then we can make a scatter plot in the first layer with "average_scovile" on the x-axis and "unfished_count" on the y-axis. We will also add a geom_smooth in the second layer to analyse the trend.
Here we calculate the "unfinished_count" variable from "episodes.csv" and the "average_scovile" variable from "sauces.csv". 

