---
title: "Hot Ones Episodes"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
if (!require("pacman"))
  install.packages("pacman")
pacman::p_load(tidyverse, tidytuesdayR)

```

## Dataset

```{r}
#| label: load-dataset
#| message: false
episodes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-08-08/episodes.csv')
sauces <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-08-08/sauces.csv')
seasons <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-08-08/seasons.csv')
```

### [**Dataset Name - Hot Ones Episodes**]{.underline}

Hot Ones is a popular YouTube series which was started on 12th March 2015. The concept behind the show was to serve the guests with hot sauces which have varying heat and spiciness numbers on the scoville scale. The show often features rare and unique hot sauces from around the world, the spiciness of these sauces is measured by the scoville scale.

Best Last Dab Reactions of All Time Video - In case you are unfamiliar with the show Hot Ones

<https://youtu.be/-at3oShfXH8?si=iYmT2YOsBiBPwFap>

Hot Ones Episodes is a data set comes from the TidyTuesday repository. The Data Set that we choose contains three tables or three sets of data.

1.  The first datset gives detailed information about the episodes

    a.  It contains the data about each episode until July 2023. The data set contains 300 observations.

    b.  The fields and the explanation of the episode data set are as follows

    | Variable Name           | Description                                                       |
    |-----------------------|-------------------------------------------------|
    | season                  | The season number.                                                |
    | episode_overall         | The overall count of episode, from 1-300                          |
    | episode_season          | The count of this episode within this season                      |
    | title                   | The title of the episode                                          |
    | original_release        | The date on which the episode was originally available on YouTube |
    | guest                   | The name of the guest                                             |
    | guest_appearance_number | The number of appearances by this guest so far as of this date    |
    | finished                | Whether the guest finished trying all of the sauces               |

    : **episodes.csv**

2.  The second data set contains the data related to the sauces

    a.  Mainly it gives the description about the seasons in which the sauces were used and the Scoville scale reading (<https://en.wikipedia.org/wiki/Scoville_scale>)
    b.  It contains 210 observations
    c.  The variables used and the interpretation of the data table is as follows

    | Variable Name | Description                                                  |
    |--------------------|----------------------------------------------------|
    | season        | The season number                                            |
    | sauce_number  | The number of this sauce, from 1 (least hot) to 10 (hottest) |
    | sauce_name    | The name of the sauce                                        |
    | scoville      | The rating of the sauce in Scoville heat units.              |

    : **sauces.csv**

3.  The Third data set contains the data related to seasons

    a.  It mainly shows the episodes per season, release date and the end date of the season. The data set contains record of 21 observations.
    b.  The variables used and their descriptions are as follows

    | Variable Name    | Description                                                                                     |
    |-------------------|-----------------------------------------------------|
    | season           | The season number                                                                               |
    | episodes         | The count of episodes in this season                                                            |
    | note             | Notes about this season                                                                         |
    | original_release | The date of the first episode in this season                                                    |
    | last_release     | The date of the last episode of this season (if that episode has aired at the time of scraping) |

    : **seasons.csv**

### [**Reason to choose the Dataset**]{.underline}

The dataset represents an intriguing and distinctive topic. It is neither too small for meaningful analysis nor excessively large to hinder us from drawing conclusions. It pertains to a subject that not only our team members are familiar with but is also widely recognized. Having watched the episodes and gaining insights into the data collection process, we found it easier to connect with the dataset and analyze it more effectively.

## Questions

Q1 ) Is there any correlation between the scoville score of the sauce and the number of guests who did not finish trying all the sauces?

1.1) Find the season which has the most unfinished contestants

1.2) Compare that with the sauce score.

1.3) Is there any correlation between the average scoville score and the number of episodes for a season.

Q2 ) How does the scoville scale of the sauces used in "Hot Ones" episodes change over different seasons? Is there any correlation between the average scoville scale number of the season and the length of the seasons?

## Analysis plan

-   For the first question, we plan on creating the variable "unfinished_count" which represents the number of guests that failed to try all the sauces for each season. Then we create another variable "average_scoville" which is the average of the variable "scoville" for each season. Then we can make a scatter plot in the first layer with "average_scoville" on the x-axis and "unfinished_count" on the y-axis. We will also add a geom_smooth in the second layer to analyse the trend. Here we calculate the "unfinished_count" variable from "episodes.csv" and the "average_scoville" variable from "sauces.csv".

-   For the second question, we plan on extracting the average scoville scale data of the sauces that appeared in each of the seasons, for all of the seasons, and draw box plots (or violin plot.). We will compare compare it to a bar plot which represents the entire length of the season as the seasons length (number of episodes) were not the same for each season. By doing this we will see if seasons that had a higher average number of scoville scale data had shorter seasons (number of episodes), and seasons with a lower overall average scoville scale number had more episodes within a season.
